---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Results {#sec-results}

```{r, echo = FALSE, results = FALSE, message = FALSE, warning = FALSE}
#| label: set-chunk-options
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)

library(ggplot2)

targets::tar_load(c("ais_date_start",
                    "ais_date_end",
                    "squid_vessel_list",
                    "gridded_time_effort_by_flag"))

source("r/functions.R")
```

```{r}
# Add country names to gridded_time_effort_by_flag
gridded_time_effort_by_flag <- gridded_time_effort_by_flag |>
  dplyr::mutate(country = countrycode::countrycode(flag,
                                                   origin = "iso3c",
                                                   destination = "country.name")) |>
  dplyr::mutate(country = ifelse(is.na(country), "Unknown",country))

# Get top fishing flags across all time
n_flags <- 10
top_flags_all_time <- gridded_time_effort_by_flag |>
  dplyr::group_by(country) |>
  dplyr::summarize(fishing_kw_hours = sum(fishing_kw_hours, na.rm = TRUE)) |>
  dplyr::ungroup()  |>
  dplyr::slice_max(order_by = fishing_kw_hours, n = n_flags) |>
  dplyr::select(country) |>
  dplyr::mutate(top = TRUE)

gridded_time_effort_by_flag <- gridded_time_effort_by_flag  |>
  dplyr::left_join(top_flags_all_time, by = "country") |>
  dplyr::mutate(country_plotting = ifelse(is.na(top), "Other",country) |>
                  forcats::fct_relevel(top_flags_all_time$country))
```


## Fishery summary

```{r}
vessel_summary <- squid_vessel_list |>
  dplyr::group_by(flag) |>
  dplyr::summarize(`Total engine power (kW)` = sum(engine_power_kw,na.rm = TRUE),
                   `Number unique vessels` = dplyr::n_distinct(ssvid)) |>
  dplyr::ungroup() |>
  dplyr::mutate(country = countrycode::countrycode(flag,
                                                   origin = "iso3c",
                                                   destination = "country.name")) |>
  dplyr::mutate(country = ifelse(is.na(country), "Unknown",country)) |>
  dplyr::select(-flag) 
```


Our analysis includes `r prettyNum(nrow(squid_vessel_list), big.mark = ",")` squid vessels spread across `r length(unique(vessel_summary$country))` flags. For each flag, we can look at the total number of unique vessels and total engine power (kW) (@fig-vessel-summary-by-flag). China dominates the fleet.

```{r}
#| label: fig-vessel-summary-by-flag
#| fig-cap: Squid jigger vessel summary by flag, showing the total number of unique vessels and total engine power (kW)
#| fig-height: 10

# Arrange plot by the number of unique vessels from each flag
plot_order <- vessel_summary |>
  dplyr::arrange(`Number unique vessels`) |>
  dplyr::pull(country)

vessel_summary |>
  tidyr::pivot_longer(cols = -country,
                      names_to = "metric",
                      values_to = "value") |>
  ggplot(aes(x = value,
             y = forcats::fct_relevel(country,plot_order))) +
  geom_bar(stat = "identity") +
  facet_wrap(~metric,
             scales = "free_x",
             ncol = 2) +
  labs(x = "",
       y = "") +
  theme_plot() +
  scale_x_continuous(labels = scales::comma)

```

Next we can look at at a map of squid fishing effort (@fig-effort-map), aggregating effort across effort and time for each pixel across the entire processed time series (`r stringr::str_remove_all(ais_date_start,'\'')` through `r stringr::str_remove_all(ais_date_end,'\'')`). EEZ boundaries from [Marine Regions v12](https://www.marineregions.org/downloads.php) are shown in orange (@marineregionsv12).

```{r}
#| label: fig-effort-map
#| fig-cap: Map of squid fishing effort from 2016 through 2024, aggregating effort across effort and time for each pixel. EEZ boundaries from Marine Regions v12 are shown in orange.

land <- rnaturalearth::ne_countries(returnclass = "sf")

# Load Marine Regions v12 EEZ
eez <- sf::st_read("/home/emlab/data/marine-regions-eez-v12/World_EEZ_v12_20231025_LR/eez_v12_lowres.shp",quiet = TRUE)

ggplot()  +
  geom_tile(data = gridded_time_effort_by_flag |>
              dplyr::group_by(lat_bin,
                              lon_bin) |>
              dplyr::summarize(fishing_kw_hours = sum(fishing_kw_hours)) |>
              dplyr::ungroup(),
            aes(fill = fishing_kw_hours,
                x = lon_bin,
                y = lat_bin)) +
  geom_sf(data = eez,
          fill = NA,
          color = "orange") +
  geom_sf(data = land,
          color = "#374a6d",
          fill = "#374a6d") +
  scale_fill_viridis_c("Fishing effort (kwW-hours)",
                       trans = "log10",
                       option = "mako",
                       na.value = "black") +
  labs(x = '',
       y= '') +
  theme_map() +
  guides(fill = guide_colorbar(title.position = "top",
                               title.hjust = 0.5,
                               frame.colour = "black",
                               ticks.colour = "black",
                               barwidth = unit(12, "cm"),
                               barheight = unit(1, "cm"))) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
```

We can also look at these maps, broken apart by flag and focusing on the top `r n_flags` flags (in terms of total effort over the entire time series).

```{r}
purrr::map(top_flags_all_time$country,
           function(country_tmp){
             ggplot()  +
  geom_tile(data = gridded_time_effort_by_flag |>
              dplyr::filter(country == country_tmp) |>
              dplyr::group_by(lat_bin,
                              lon_bin) |>
              dplyr::summarize(fishing_kw_hours = sum(fishing_kw_hours)) |>
              dplyr::ungroup(),
            aes(fill = fishing_kw_hours,
                x = lon_bin,
                y = lat_bin)) +
  geom_sf(data = eez,
          fill = NA,
          color = "orange") +
  geom_sf(data = land,
          color = "#374a6d",
          fill = "#374a6d") +
  scale_fill_viridis_c("Fishing effort (kwW-hours)",
                       trans = "log10",
                       option = "mako",
                       na.value = "black") +
  labs(x = '',
       y= '',
       title = country_tmp) +
  theme_map() +
  guides(fill = guide_colorbar(title.position = "top",
                               title.hjust = 0.5,
                               frame.colour = "black",
                               ticks.colour = "black",
                               barwidth = unit(12, "cm"),
                               barheight = unit(1, "cm"))) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
           })
```

We can also look at a time series of total monthly global effort by fishing flag over time (@fig-effort-time-series). The top top `r n_flags` flags flags are shown, with all other flags aggregated into "Other".

```{r}
#| label: fig-effort-time-series
#| fig-cap: Total monthly global effort by fishing flag over time. The top 10 flags are shown by color, with all other flags aggregated into "Other".
#| fig-height: 10

gridded_time_effort_by_flag|>
  dplyr::group_by(month,
                  country_plotting) |>
  dplyr::summarize(fishing_kw_hours = sum(fishing_kw_hours)) |>
  dplyr::ungroup() |>
  ggplot(aes(x = month, y= fishing_kw_hours, fill = country_plotting)) +
 geom_line() +
  labs(x = "",
       y = "Global fishing effort by flag (kW-hours)")+
  scale_y_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  theme_plot() +
  facet_grid(country_plotting~.,
             scales = "free_y")
```

