---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Results {#sec-results}

```{r, echo = FALSE, results = FALSE, message = FALSE, warning = FALSE}
#| label: set-chunk-options
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)

library(ggplot2)

targets::tar_load(c("squid_vessel_list",
                    "joined_dataset_ais",
                    "joined_dataset_viirs"))

source("r/functions.R")
```

```{r}
# Add country names to joined_dataset_ais
joined_dataset_ais <- joined_dataset_ais |>
  dplyr::mutate(country = countrycode::countrycode(flag,
                                                   origin = "iso3c",
                                                   destination = "country.name")) |>
  # If country name is missing, call it Unknown
  dplyr::mutate(country = ifelse(is.na(country), "Unknown",country))

# Get top fishing flags across all time
n_flags <- 10
top_flags_all_time <- joined_dataset_ais |>
  dplyr::group_by(country) |>
  dplyr::summarize(fishing_kw_hours = sum(fishing_kw_hours, na.rm = TRUE)) |>
  dplyr::ungroup()  |>
  dplyr::slice_max(order_by = fishing_kw_hours, n = n_flags) |>
  dplyr::mutate(top = TRUE)

joined_dataset_ais <- joined_dataset_ais  |>
  dplyr::left_join(top_flags_all_time |>
                     dplyr::select(country,top), by = "country") |>
  dplyr::mutate(country_plotting = ifelse(is.na(top), "Other",country) |>
                  forcats::fct_relevel(top_flags_all_time$country))

other_effort_all_time <- joined_dataset_ais |>
  dplyr::filter(country_plotting == "Other") |>
  dplyr::summarize(fishing_kw_hours = sum(fishing_kw_hours, na.rm=TRUE)) |>
  dplyr::mutate(country = "Other",
                top = TRUE)

# Include other in our top flags, for plotting
top_flags_all_time <- top_flags_all_time |>
  dplyr::bind_rows(other_effort_all_time) |>
  dplyr::arrange(-fishing_kw_hours)
```

## Fishery summary

### Fleet description

```{r}
vessel_summary <- squid_vessel_list |>
  dplyr::group_by(flag) |>
  dplyr::summarize(`Total engine power (kW)` = sum(engine_power_kw,na.rm = TRUE),
                   `Number unique vessels` = dplyr::n_distinct(ssvid)) |>
  dplyr::ungroup() |>
  dplyr::mutate(country = countrycode::countrycode(flag,
                                                   origin = "iso3c",
                                                   destination = "country.name")) |>
  dplyr::mutate(country = ifelse(is.na(country), "Unknown",country)) |>
  dplyr::select(-flag) 
```

Our analysis includes `r prettyNum(nrow(squid_vessel_list), big.mark = ",")` squid vessels spread across `r length(unique(vessel_summary$country))` flags. For each flag, we look at the total engine power (kW) across the fleet and total number of unique vessels and  (@fig-vessel-summary-by-flag). China dominates the fleet, with Taiwan a fairly distant second.

```{r}
#| label: fig-vessel-summary-by-flag
#| fig-cap: Squid jigger vessel summary by flag, showing the total engine power (kW) and total number of unique vessels.
#| fig-height: 7

# Arrange plot by the number of unique vessels from each flag
plot_order <- vessel_summary |>
  dplyr::arrange(`Total engine power (kW)`) |>
  dplyr::pull(country)

vessel_summary |>
  tidyr::pivot_longer(cols = -country,
                      names_to = "metric",
                      values_to = "value") |>
  dplyr::mutate(metric = forcats::fct_relevel(metric,"Total engine power (kW)")) |>
  ggplot(aes(x = value,
             y = forcats::fct_relevel(country,plot_order))) +
  geom_bar(stat = "identity") +
  facet_wrap(~metric,
             scales = "free_x",
             ncol = 2) +
  labs(x = "",
       y = "") +
  theme_plot() +
  scale_x_continuous(labels = scales::comma)

```

Next we look at total global fishing effort from 2016 through August 2024, by flag (@fig-effort-bar-plot). The top `r n_flags` flags are shown individually, with all other flags aggregated into "Other". Again, China dominates the fishing effort, with Taiwan a fairly distant second.

```{r}
#| label: fig-effort-bar-plot
#| fig-cap: Total fishing effort, by flag, from 2016 through August 2024. The top 10 flags are shown, with all other flags aggregated into "Other".

joined_dataset_ais|>
  dplyr::group_by(country_plotting) |>
  dplyr::summarize(fishing_kw_hours = sum(fishing_kw_hours,na.rm = TRUE)) |>
  dplyr::ungroup() |>
  ggplot(aes(x = fishing_kw_hours, y= reorder(country_plotting,fishing_kw_hours))) +
  geom_bar(stat = "identity")+
  labs(x = "Total fishing effort from 2016 through August 2024 (kW-hours)",
       y = "")+
  scale_x_continuous(labels = scales::unit_format(unit = "B", scale = 1e-9)) +
  theme_plot()
```

## Fishing effort

Next we look at a time series of total monthly global effort by fishing flag over time (@fig-effort-time-series). The top top `r n_flags` flags are shown individually, with all other flags aggregated into "Other".

```{r}
#| label: fig-effort-time-series
#| fig-cap: Monthly fishing effort by flag from 2016 through August 2024. The top 10 flags are shown individually, with all other flags aggregated into "Other".
#| fig-height: 7

joined_dataset_ais|>
  dplyr::group_by(month,
                  country_plotting) |>
  dplyr::mutate(country_plotting = forcats::fct_relevel(country_plotting,
                                                        top_flags_all_time$country)) |>
  dplyr::summarize(fishing_kw_hours = sum(fishing_kw_hours)) |>
  dplyr::ungroup() |>
  ggplot(aes(x = month, y= fishing_kw_hours, fill = country_plotting)) +
 geom_line() +
  labs(x = "",
       y = "Monthly fishing effort by flag (kW-hours)")+
  scale_y_continuous(labels = scales::unit_format(unit = "M", scale = 1e-6)) +
  theme_plot() +
  facet_grid(country_plotting~.,
             scales = "free_y")
```

Next we can look at at a map of AIS-based squid fishing effort (@fig-effort-map), aggregating effort across effort and flags and time for each pixel across the entire processed time series. EEZ boundaries from [Marine Regions v12](https://www.marineregions.org/downloads.php) are shown in orange (@marineregionsv12).

```{r}
#| label: fig-effort-map
#| fig-cap: Map of squid jigger fishing effort from 2016 through August 2024, aggregating effort across effort, flags, and time for each pixel. EEZ boundaries from Marine Regions V12 are shown in orange.

land <- rnaturalearth::ne_countries(returnclass = "sf")

# Load Marine Regions v12 EEZ
eez <- sf::st_read("/home/emlab/data/marine-regions-eez-v12/World_EEZ_v12_20231025_LR/eez_v12_lowres.shp",quiet = TRUE)

ggplot()  +
  geom_tile(data = joined_dataset_ais |>
              dplyr::group_by(lat_bin,
                              lon_bin) |>
              dplyr::summarize(fishing_kw_hours = sum(fishing_kw_hours)) |>
              dplyr::ungroup(),
            aes(fill = fishing_kw_hours,
                x = lon_bin,
                y = lat_bin)) +
  geom_sf(data = eez,
          fill = NA,
          color = "orange") +
  geom_sf(data = land,
          color = "#374a6d",
          fill = "#374a6d") +
  scale_fill_viridis_c("Fishing effort (kwW-hours)",
                       trans = "log10",
                       option = "mako",
                       na.value = "black") +
  labs(x = '',
       y= '') +
  theme_map() +
  guides(fill = guide_colorbar(title.position = "top",
                               title.hjust = 0.5,
                               frame.colour = "black",
                               ticks.colour = "black",
                               barwidth = unit(12, "cm"),
                               barheight = unit(1, "cm"))) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
```

We can also look at these maps, broken apart by flag, and still aggregating squid jigger fishing effort from 2016 through August 2024, aggregating effort across time for each pixel. The top `r n_flags` flags are shown individually, with all other flags aggregated into "Other".

```{r}
purrr::map(top_flags_all_time$country,
           function(country_tmp){
             ggplot()  +
  geom_tile(data = joined_dataset_ais |>
              dplyr::filter(country_plotting == country_tmp) |>
              dplyr::group_by(lat_bin,
                              lon_bin) |>
              dplyr::summarize(fishing_kw_hours = sum(fishing_kw_hours)) |>
              dplyr::ungroup(),
            aes(fill = fishing_kw_hours,
                x = lon_bin,
                y = lat_bin)) +
  geom_sf(data = eez,
          fill = NA,
          color = "orange") +
  geom_sf(data = land,
          color = "#374a6d",
          fill = "#374a6d") +
  scale_fill_viridis_c("Fishing effort (kwW-hours)",
                       trans = "log10",
                       option = "mako",
                       na.value = "black") +
  labs(x = '',
       y= '',
       title = glue::glue("{country_tmp}\n")) +
  theme_map() +
  guides(fill = guide_colorbar(title.position = "top",
                               title.hjust = 0.5,
                               frame.colour = "black",
                               ticks.colour = "black",
                               barwidth = unit(12, "cm"),
                               barheight = unit(1, "cm"))) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
           })
```

## VIIRS detections

First we look at a time series of monthly VIIRS detections globally (@fig-viirs-time-series). These can generally be thought of as representing light-luring squid vessels. Again we can see a clear seasonal pattern.

```{r}
#| label: fig-viirs-time-series
#| fig-cap: Monthly VIIRS detections.

joined_dataset_viirs|>
  dplyr::group_by(month) |>
  dplyr::summarize(viirs_detections = sum(viirs_detections)) |>
  dplyr::ungroup() |>
  ggplot(aes(x = month, y= viirs_detections)) +
 geom_line() +
  labs(x = "",
       y = "Monthly VIIRS detections")+
  theme_plot() +
  scale_y_continuous(labels = scales::comma)
```

Next we can look at at a map of these VIIRS detections (@fig-viirs-map), aggregating across time for each pixel across the entire processed time series (January 2017 through December 2021). The general spatial patterns seem to be consistent with those seen using AIS-based fishing effort data (@fig-effort-map).

```{r}
#| label: fig-viirs-map
#| fig-cap: Map of VIIRS detections, aggregating across time for each pixel. EEZ boundaries from Marine Regions V12 are shown in orange.

ggplot()  +
  geom_tile(data = joined_dataset_viirs |>
              dplyr::group_by(lat_bin,
                              lon_bin) |>
              dplyr::summarize(viirs_detections = sum(viirs_detections)) |>
              dplyr::ungroup(),
            aes(fill = viirs_detections,
                x = lon_bin,
                y = lat_bin)) +
  geom_sf(data = eez,
          fill = NA,
          color = "orange") +
  geom_sf(data = land,
          color = "#374a6d",
          fill = "#374a6d") +
  scale_fill_viridis_c("VIIRS detections",
                       trans = "log10",
                       option = "mako",
                       na.value = "black") +
  labs(x = '',
       y= '') +
  theme_map() +
  guides(fill = guide_colorbar(title.position = "top",
                               title.hjust = 0.5,
                               frame.colour = "black",
                               ticks.colour = "black",
                               barwidth = unit(12, "cm"),
                               barheight = unit(1, "cm"))) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
```