---
title: "Analysis"
author: "Kaiwen Wang"
date: "2024-11-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE)
```


```{r, include= FALSE}
library(tidyverse)
library(fixest)
library(modelsummary)
library(marginaleffects)
#targets::tar_make(quiet = FALSE)
```


```{r, eval=FALSE}
# Initialize targets 

targets_data_directory = "C:\\Users\\kevin\\Nextcloud\\emLab\\projects\\current-projects\\squid-climate-change\\data"

targets::tar_config_set(store = glue::glue("{targets_data_directory}/_targets"))

targets::tar_load(c(#"gridded_time_effort_by_flag",
                    #"sst_data_aggregated",
                    #"oceanic_nino_index_data",
                    #"pixels_eez_with_info",
                    #"analysis_scope_lon",
                    #"analysis_scope_lat",
                    "joined_dataset_ais",
                    "sst_cc_forecast_data"))
```

```{r, eval=FALSE}
ais_data = 
  joined_dataset_ais %>% 
  mutate(bin = str_c(lon_bin, ",", lat_bin)) %>% 
  mutate(month_idx = month(month),
         year = year(month),) %>% 
  mutate(season = case_when(
    month_idx %in% c(9,10,11) ~ "spring",
    month_idx %in% c(12,1,2) ~ "summer",
    month_idx %in% c(3,4,5) ~ "fall",
    month_idx %in% c(6,7,8) ~ "winter",),) %>% 
  mutate(lat_bin_wgi = floor(lat_bin),
         lon_bin_wgi = floor(lon_bin),
         bin_wgi = str_c(lon_bin_wgi,",", lat_bin_wgi))
```


# Prepare the ais_data by filtering and computing necessary statistics

```{r}
# Function to prepare data for modeling based on latitude filter
prepare_model_data <- function(data, latitude_filter) {
  data %>%
    filter(lat_bin >= latitude_filter[1],
           lat_bin < latitude_filter[2]) %>%
    group_by(month) %>%
    mutate(
      avg_sst_for_month = mean(sst_deg_c_mean),
      sst_deviation = sst_deg_c_mean - avg_sst_for_month,
      sst_deviation_squared = sst_deviation^2,
      sst_squared = sst_deg_c_mean^2,
      scaled_fishing_effort = fishing_kw_hours / 1000
    ) %>%
    ungroup()
}

```


# Fit a Poisson model using fepois from the fixest package

```{r}
# Function to fit different models
fit_model <- function(data, model_type = "feols") {
  if (model_type == "feols") {
    feols(scaled_fishing_effort ~ sst_deviation:factor(month_idx) +
          sst_deviation_squared:factor(month_idx) 
          | month + bin_wgi,
          data = data,
          cluster = ~ bin, nthreads = 20)
  } else {
    fepois(scaled_fishing_effort ~ sst_deviation:factor(month_idx) +
           sst_deviation_squared:factor(month_idx) 
           | month + bin_wgi,
           data = data,
           cluster = ~ bin, nthreads = 20)
  }
}

```

# Apply poission model for latitude condition less than -10
```{r}
model_data_less_10 <- prepare_model_data(ais_data, c(-Inf,-10))
poisson_model_less_10 <- fit_model(model_data_less_10, "fepois")
```

```{r}
# Plot model coefficients
coefplot(poisson_model_less_10, drop = c("sst_deviation_squared", "bin"))
```

# Apply model for latitude condition greater than -10
```{r}
model_data_greater_10 <- prepare_model_data(ais_data, c(-10,Inf))
poisson_model_greater_10 <- fit_model(model_data_greater_10, "fepois")

```

```{r}
# Plot model coefficients
coefplot(poisson_model_greater_10)
```
# Save both models

```{r}
#saveRDS(poisson_model_greater_10, "poisson_model_greater_10.rds")
#saveRDS(poisson_model_less_10, "poisson_model_less_10.rds")
```


# Rename columns in the forecast dataset for consistency
```{r}
forecast_data_sst = sst_cc_forecast_data %>%
  rename(lon_bin_wgi = lon_bin,
         lat_bin_wgi = lat_bin,
         sst_deg_c_mean_pred = sst_deg_c_mean,
         month_idx = month_number)
```


# Define a function to handle forecast data manipulation and prediction

```{r}
prepare_model_used_data = function(fixest_model, obs_orgin){
  #Identify observations used by the model
  obsRemoved = fixest_model$obs_selection$obsRemoved
  if (length(obsRemoved) > 0) {
    model_used_data <-
    obs_orgin[obsRemoved, ]
  } else{
    model_used_data = obs_orgin
  }
  return(model_used_data)
}
```



```{r}
prepare_scenario_data <- function(model_used_data, prepared_forecast_data, scenario_label, time_label) {
    ais_data_model_forecast <- 
    model_used_data %>%
    left_join(prepared_forecast_data %>%
                filter(scenario == scenario_label,
                       time_period == time_label)) %>%
    #use original month mean
    #replace predictors with forecasts
    mutate(original_sst = sst_deg_c_mean, 
           forecast_sst = sst_deg_c_mean_pred, 
           sst_deg_c_mean = forecast_sst, 
           sst_deviation = sst_deg_c_mean - avg_sst_for_month,
           sst_deviation_squared = sst_deviation^2) %>% 
  return(ais_data_model_forecast)
}
```

# Apply the function for different models


```{r}
# Function to process and combine forecasts for multiple scenarios
process_and_combine_forecasts <- function(model, model_used_data, prepared_forecast_data, scenarios, time_periods) {
  # Predictions for the fitted model on the same data
  fitted_predictions <- predict(model, newdata = model_used_data)
  
  # Start the result dataframe with fitted predictions
  results <- model_used_data %>%
    mutate(fitted = fitted_predictions,
           original_sst = sst_deg_c_mean) %>%     
    select(bin, month, fitted, original_sst, scaled_fishing_effort,
           lon_bin, lat_bin)
  
  # Loop through each scenario to add forecast predictions and SST to the results
  for (i in seq_along(scenarios)) {
    scenario <- scenarios[i]
    time_period <- time_periods[i]
    
    # Prepare scenario-specific forecast data
    forecast_scenario_data <- prepare_scenario_data(model_used_data, prepared_forecast_data, scenario, time_period)
    
    # Predict using the model for the current scenario
    scenario_prediction <- predict(model, newdata = forecast_scenario_data)
    forecast_sst = forecast_scenario_data$forecast_sst
    
    # Add the scenario predictions as a new column to the results dataframe
    results <- results %>%
      mutate("{scenario}_prediction" := scenario_prediction,
             "{scenario}_forecast_sst" := forecast_sst)
  }
  
  return(results)
}

```


## Load saved local data and model

```{r}
ais_data = read_rds("ais_data.rds")
poisson_model_greater_10 = read_rds("poisson_model_greater_10.rds")
poisson_model_less_10 = read_rds("poisson_model_less_10.rds")
```



# Model prediction for Poisson_model_greater_than_10

```{r}

scenarios <- 
  c("SSP1-2.6", "SSP2-4.5", "SSP3-7.0", "SSP5-8.5")
time_periods <- 
  c("Near Term (2021-2040)", "Near Term (2021-2040)", "Near Term (2021-2040)", "Near Term (2021-2040)")

# Identify observations used by the model
model_used_data_greater_10 =
  prepare_model_used_data(poisson_model_greater_10,
                          model_data_greater_10)

combined_results_greater_10 <- 
  process_and_combine_forecasts(poisson_model_greater_10, model_used_data_greater_10, forecast_data_sst, scenarios, time_periods)

```
```{r}
combined_results_greater_10_increase =
  combined_results_greater_10 %>% 
  filter(`SSP1-2.6_prediction` > fitted)
```

# Model prediction for Poisson_model_less_than_10
```{r}

scenarios <- 
  c("SSP1-2.6", "SSP2-4.5", "SSP3-7.0", "SSP5-8.5")
time_periods <- 
  c("Near Term (2021-2040)", "Near Term (2021-2040)", "Near Term (2021-2040)", "Near Term (2021-2040)")

# Identify observations used by the model
model_used_data_less_10 =
  prepare_model_used_data(poisson_model_less_10,
                          model_data_less_10)

combined_results_less_10 <- 
  process_and_combine_forecasts(poisson_model_less_10, model_used_data_less_10, forecast_data_sst, scenarios, time_periods)

```

```{r}
combined_results_less_10_increase =
  combined_results_less_10 %>% 
  filter(`SSP1-2.6_prediction` > fitted)
```


## Save predicted results

```{r}
combined_results_poisson = 
  bind_rows(combined_results_less_10,
            combined_results_greater_10)

#saveRDS(ais_data,"ais_data.rds")

saveRDS(combined_results_poisson, "combined_results_poisson.rds")
```




